Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-05-01 PAPER.tex
edited boot methods plot and text to be compatible to 'all bl cross-multiplications' run

Given our new understanding of the sensitivity of bootstraps to the number of elements sampled, we have removed the second 
bootstrapping step along time entirely and now simply bootstrap over baselines. Power spectrum $2\sigma$ upper limits with this bootstrapping change for fringe-rate filtered noise are shown in Figure 
\ref{fig:data_errors}. The estimates are uniformly weighted in order to disentangle the effects of bootstrapping from signal loss. As 
shown in the figure, when more elements are drawn for each bootstrap than the number of 
independent samples (by over-sampling elements along the time axis), repeated values begin to crop up and the apparent variation between bootstraps drops, resulting in limits (gray) below the predicted noise level (green). Using the revised bootstrapping method, where bootstrapping only occurs over the baseline axis, the limits (black) are shown to agree with the analytic prediction for noise. Figure \ref{fig:data_errors} implies that power spectrum upper limits were originally under-estimated by a factor of $\sim$$2$-$3$ in mK$^{2}$.

Finally, one last change from the \citetalias{ali_et_al2015} method is that our power spectrum points (previously computed as the mean of all bootstraps), are now computed as the power spectrum estimate resulting from not bootstrapping at all. More specifically, we compute one estimate without sampling over the baseline axis (all baselines are used), and this estimate is propagated through our signal loss computation. The difference between taking the mean of the bootstrapped values and using the estimate from the no-bootstrapping case is small, but doing the latter ensures that we are forming results that reflect the estimate preferred by all our data.

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0cm 0.3cm 0.3cm},width=\columnwidth]{plots/noise_errors.pdf}
	\caption{$2\sigma$ power spectrum upper limits for a noise simulation (computed via Equation \eqref{eq:noise} using PAPER-64 observing parameters) using two different bootstrapping 
methods. The noise is fringe-rate filtered and a weighting matrix of $\textbf{I}$ (uniform-weighted) is used in order to disentangle the 
effects of bootstrapping from signal loss. The bootstrapping method used in \citetalias{ali_et_al2015} is shown in gray, where bootstrapping occurs along both the baseline and time axes. This under-estimates errors by sampling more values than independent ones in the dataset (fringe-rate filtering reduces the number of independent samples along time). We use the method illustrated by the black curve in our updated analysis, where bootstrapping only occurs along the baseline axis. We find that these revised limits agree with the $2\sigma$ analytic prediction for noise (green).}
	\label{fig:data_errors}
\end{figure}

@@

\begin{itemize}
\item Signal loss, previously found to be $<2\%$ in \citetalias{ali_et_al2015}, was under-estimated by a factor of $>$$1000$ for empirically estimated inverse 
covariance weighting. For our new analysis, we use a regularized covariance weighting method to minimize loss (Section 
\ref{sec:Weight}). However, because our revised weighting method is not as aggressive as the former, our results are still a 
factor of $\sim10$ in mK higher than previous limits. Under-estimated signal loss therefore represents the bulk of our revision. 
This revision is similar to the re-analysis of results from the GMRT (\citealt{paciga_et_al2013}) which were also 
revised from new signal loss calculations associated with their singular value decomposition foreground filter.
\item Power spectrum errors, originally computed by bootstrapping, were under-estimated for the data by a factor of $\sim2$ in mK due to oversampling data whose effective number of independent samples was reduced from fringe-rate filtering (Section \ref{sec:Boot}). 
\item Several factors used in an analytic expression to predict the noise-level in PAPER-64 data were revised, yielding a 
decrease in predicted sensitivity level by a factor of $\sim2$ in mK (Section \ref{sec:PSSense}). We note that our sensitivity prediction is revised by a factor less than our 
power spectrum result, implying that if taken at face value, the theoretical prediction for noise in \citetalias{ali_et_al2015} was too high for its data 
points.
\end{itemize}

>>>>

Given our new understanding of the sensitivity of bootstraps to the number of elements sampled, we have removed the second 
bootstrapping step along time entirely and now simply bootstrap over the baseline axis. Power spectrum $2\sigma$ errors with this bootstrapping change for fringe-rate filtered noise are shown in Figure 
\ref{fig:data_errors}. The estimates are uniformly weighted in order to disentangle the effects of bootstrapping from signal loss. As 
shown in the figure, when more elements are drawn for each bootstrap than the number of 
independent samples (by over-sampling elements along the time axis), repeated values begin to crop up and the apparent variation between bootstraps drops, resulting in limits (gray) below the predicted noise level (green). Using the revised bootstrapping method, where bootstrapping only occurs over the baseline axis, the limits (black) are shown to agree with the analytic prediction for noise. While Figure \ref{fig:data_errors} implies that errors are under-estimated by a factor of $\sim$ $5$ in mK$^{2}$ for the noise simulation, in practice this factor is lower for the case of real data (a factor of $\sim$ $2$ in mK$^{2}$ instead), possibly due to the data being less correlated in time than the fringe-rate filtered noise in the simulation. %Figure \ref{fig:data_errors} implies that power spectrum upper limits were originally under-estimated by a factor of $\sim$ $5$ in mK$^{2}$.

In addition to no longer bootstrapping along the time axis, we have made two other changes from the \citetalias{ali_et_al2015} analysis regarding bootstrapping. The first is that we now bootstrap over baseline cross-products, instead of baselines themselves. In the previous analysis, baselines were bootstrapped prior to forming cross power spectra, and using this particular ordering of operations (bootstrapping, then cross-multiplication) yields variances that have been found to disagree with predicted errors from bootstrapping using simulations. On the contrary, bootstrapping over cross power spectra ensures that we are estimating the variance of our quantity of interest (i.e. the power spectrum). This change, while fundamental in retaining the integrity of the bootstrapping method in general, alters the resulting power spectrum errors by factors of $<2$ in practice.

Finally, the last change from the \citetalias{ali_et_al2015} method is that our power spectrum points (previously computed as the mean of all bootstraps), are now computed as the power spectrum estimate resulting from not bootstrapping at all. More specifically, we compute one estimate without sampling, and this estimate is propagated through our signal loss computation (this estimate is $\widehat{\textbf{P}}_{x}$). The difference between taking the mean of the bootstrapped values and using the estimate from the no-bootstrapping case is small, but doing the latter ensures that we are forming results that reflect the estimate preferred by all our data.

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0cm 0.3cm 0.3cm},width=\columnwidth]{plots/noise_errors.pdf}
	\caption{$2\sigma$ power spectrum errors for a noise simulation (computed via Equation \eqref{eq:noise} using PAPER-64 observing parameters) using two different bootstrapping 
methods. The noise is fringe-rate filtered and a weighting matrix of $\textbf{I}$ (uniform-weighted) is used in order to disentangle the 
effects of bootstrapping from signal loss. The bootstrapping method used in \citetalias{ali_et_al2015} is shown in gray, where bootstrapping occurs along both the baseline and time axes. This under-estimates errors by sampling more values than independent ones in the dataset (fringe-rate filtering reduces the number of independent samples along time). We use the method illustrated by the black curve in our updated analysis, where bootstrapping only occurs along the baseline axis. We find that these revised limits agree with the $2\sigma$ analytic prediction for noise (green).}
	\label{fig:data_errors}
\end{figure}

@@

\begin{itemize}
\item Signal loss, previously found to be $<2\%$ in \citetalias{ali_et_al2015}, was under-estimated by a factor of $>$$1000$ for empirically estimated inverse 
covariance weighting. For our new analysis, we use a regularized covariance weighting method to minimize loss (Section 
\ref{sec:Weight}). However, because our revised weighting method is not as aggressive as the former, our results are still a 
factor of $\sim10$ in mK higher than previous limits. Under-estimated signal loss therefore represents the bulk of our revision. 
This revision is similar to the re-analysis of results from the GMRT (\citealt{paciga_et_al2013}) which were also 
revised from new signal loss calculations associated with their singular value decomposition foreground filter.
\item Power spectrum errors, originally computed by bootstrapping, were under-estimated for the data by a factor of $\sim2$ in mK due to oversampling data whose effective number of independent samples was reduced from fringe-rate filtering (Section \ref{sec:Boot}). 
\item Several factors used in an analytic expression to predict the noise-level in PAPER-64 data were revised, yielding a 
decrease in predicted sensitivity level by a factor of $\sim3$ in mK (Section \ref{sec:PSSense}). We note that our sensitivity prediction is revised by a factor less than our 
power spectrum result, implying that if taken at face value, the theoretical prediction for noise in \citetalias{ali_et_al2015} was too high for its data 
points.
\end{itemize}
